Exploratory Analysis Script

### 2024-03-22 code/analysis-code/introanalysis-v1.qmd
Debugged Analysis Code
After Zane's help

```{r}
#Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear_processed.rds")
#load data
ML_linear <- readRDS(data_location)
```

>>>>

```{r}
#Path to summary data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","ML_linear_final.rds")
#load data
ML_linear_final <- readRDS(data_location)
```

### 2024-03-22 code/eda-code/eda-v1.qmd
Debugged Analysis Code
After Zane's help

```{r}
# Change observations from Portuguese to English
ML_summary <- ML_summary %>%
  mutate(
    ModifiedColor = recode_factor(ModifiedColor, "AfroDescent" = "Afro-Descent", "EuroDescent" = "Euro-Descent"),
    ModifiedNationality = recode_factor(ModifiedNationality, "European" = "European", "LatinAmerican" = "Latin American", "MiddleEastern" = "Middle Eastern", "Brazilian" = "Brazilian"),
    Color = recode_factor(Color, "Preta" = "Black", "Parda" = "Mixed Race", "Branca" = "White"),
    Nationality = recode_factor(Nationality, "Alema" = "German", "Argentina" = "Argentine", "Austriaca" = "Austrian", "Brasileira" = "Brazilian", "Espanhola" = "Spanish", "Francesa" = "French", "Italiana" = "Italian", "Paraguaya" = "Paraguayan", "Polaca" = "Polish", "Portuguesa" = "Portuguese", "Rumania" = "Romanian", "Russa" = "Russian", "Suiça" = "Swiss", "Siria" = "Syrian", "Uruguaya" = "Uruguayan"),
    Birth = recode_factor(Birth, "aborto" = "Abortion", "intervencionista" = "Interventionist", "natural" = "Natural", "operatório" = "Operative"),
    MaternalOutcome = recode_factor(MaternalOutcome, "alta" = "Discharged", "morta" = "Death", "transferência" = "Hospital transferal"),
    FetalOutcome = recode_factor(FetalOutcome, "vivo" = "Live Birth", "morto" = "Stillbirth or Neonatal Death")
    )
```

@@

# Change observations from Portuguese to English
ML_linear <- ML_linear %>%
  mutate(
    ModifiedColor = recode_factor(ModifiedColor, "AfroDescent" = "Afro-Descent", "EuroDescent" = "Euro-Descent"),
    ModifiedNationality = recode_factor(ModifiedNationality, "European" = "European", "LatinAmerican" = "Latin American", "MiddleEastern" = "Middle Eastern", "Brazilian" = "Brazilian"),
    Color = recode_factor(Color, "Preta" = "Black", "Parda" = "Mixed Race", "Branca" = "White"),
    Nationality = recode_factor(Nationality, "Alema" = "German", "Argentina" = "Argentine", "Austriaca" = "Austrian", "Brasileira" = "Brazilian", "Espanhola" = "Spanish", "Francesa" = "French", "Italiana" = "Italian", "Paraguaya" = "Paraguayan", "Polaca" = "Polish", "Portuguesa" = "Portuguese", "Rumania" = "Romanian", "Russa" = "Russian", "Suiça" = "Swiss", "Siria" = "Syrian", "Uruguaya" = "Uruguayan"),
    Birth = recode_factor(Birth, "aborto" = "Abortion", "intervencionista" = "Interventionist", "natural" = "Natural", "operatório" = "Operative"),
    MaternalOutcome = recode_factor(MaternalOutcome, "alta" = "Discharged", "morta" = "Death", "transferência" = "Hospital transferal"),
    FetalOutcome = recode_factor(FetalOutcome, "vivo" = "Live Birth", "morto" = "Stillbirth or Neonatal Death")
    )
str(ML_linear)
```

@@

Here is the previous code I worked on before learning about `gtsummary`.

```{r}
# Don't use this code in future
# Summary stats of continuous variables
summary_stats <- ML_summary %>%
  summarize(
    mean_age = mean(Age, na.rm = TRUE),
    sd_age = sd(Age, na.rm = TRUE),
    mean_weight = mean(Weightgrams, na.rm = TRUE),
    sd_weight = sd(Weightgrams, na.rm = TRUE),
    mean_length = mean(Lengthcentimeters, na.rm = TRUE),
    sd_length = sd(Lengthcentimeters, na.rm = TRUE)
  )

print(summary_stats)

# Make into a table
# Create dataframe
summary_stats1 <- data.frame(
  Variable = c("Maternal Age", "Infant birthweight (grams)", "Infant birth length (cm)"),
  Mean = c(25.30435, 3087.341, 48.33493),
  StandardDev = c(5.759474, 566.5855, 3.92719)
)

# Create a gt table
summary_stats_table <- summary_stats1 %>%
  gt() %>%
  tab_header(
    title = "Summary Statistics",
    subtitle = "Mean Values and Standard Deviations of Continuous Variables"
  ) %>%
  cols_label(
    Variable = "Variable",
    Mean = "Mean",
    StandardDev = "Standard Deviation"
  )

# Print the table
print(summary_stats_table)

#Save table
#saveRDS(summary_stats_table, file = here::here("results", "tables", "summary_stats_table.rds"))

# Summary stats of categorical variables
# Calculate frequencies of color category (branca, parda, preta)
color_freq <- ML_summary %>%
  filter(!is.na(Color)) %>% # Exclude rows with NA in color variable
  count(Color, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(color_freq)

# Calculate frequencies of color category (branca, parda, preta)
color_freq2 <- ML_summary %>%
  filter(!is.na(ModifiedColor)) %>% # Exclude rows with NA in color variable
  count(ModifiedColor, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(color_freq2)

# Calculate frequencies of Status category
status_freq <- ML_summary %>%
  filter(!is.na(Status)) %>% # Exclude rows with NA in status variable
  count(Status, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(status_freq)

# Calculate frequencies of ModifiedStatus category
status_freq2 <- ML_summary %>%
  filter(!is.na(ModifiedStatus)) %>% # Exclude rows with NA in status variable
  count(ModifiedStatus, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(status_freq2)

# Calculate frequencies of Nationality category
nationality_freq <- ML_summary %>%
  filter(!is.na(Nationality)) %>% # Exclude rows with NA in nationality variable
  count(Nationality, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(nationality_freq)

# Calculate frequencies of ModifiedNationality category
nationality_freq2 <- ML_summary %>%
  filter(!is.na(ModifiedNationality)) %>% # Exclude rows with NA in nationality variable
  count(ModifiedNationality, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(nationality_freq2)

# Calculate frequencies of MaternalOutcome category
maternal_freq <- ML_summary %>%
  filter(!is.na(MaternalOutcome)) %>% # Exclude rows with NA in MaternalOutcome variable
  count(MaternalOutcome, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(maternal_freq)

# Calculate frequencies of Sex category
sex_freq <- ML_summary %>%
  filter(!is.na(Sex)) %>% # Exclude rows with NA in Sex variable
  count(Sex, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(sex_freq)

# Calculate frequencies of birth category with abortion
birth_freq <- ML_summary %>%
  filter(!is.na(Birth)) %>% # Exclude rows with NA in birth variable
  count(Birth, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(birth_freq)

# Calculate frequencies of FetalOutcome category excluding abortion
fetal_freq <- ML_summary %>%
  filter(!is.na(FetalOutcome)) %>% # Exclude rows with NA in FetalOutcome variable
  count(FetalOutcome, na.rm = TRUE) %>% # Calculate frequencies while excluding NAs
  filter(FetalOutcome != "abortion") %>% # Exclude rows with abortion
  mutate(Total_Sample_Size = sum(n)) %>% # Calculate total sample size 
  mutate(Percent = n/sum(n)*100) # Calculate percentage of each category
print(fetal_freq)
```

>>>>

```{r}
# Change observations from Portuguese to English
ML_summary <- ML_summary %>%
  mutate(
    ModifiedColor = recode_factor(ModifiedColor, "AfroDescent" = "Afro-Descent", "EuroDescent" = "Euro-Descent"),
    ModifiedNationality = recode_factor(ModifiedNationality, "European" = "European", "LatinAmerican" = "Latin American", "MiddleEastern" = "Middle Eastern", "Brazilian" = "Brazilian"),
    Color = recode_factor(Color, "Preta" = "Black", "Parda" = "Mixed Race", "Branca" = "White"),
    Nationality = recode_factor(Nationality, "Alema" = "German", "Argentina" = "Argentine", "Austriaca" = "Austrian", "Brasileira" = "Brazilian", "Espanhola" = "Spanish", "Francesa" = "French", "Italiana" = "Italian", "Paraguaya" = "Paraguayan", "Polaca" = "Polish", "Portuguesa" = "Portuguese", "Rumania" = "Romanian", "Russa" = "Russian", "Suiça" = "Swiss", "Siria" = "Syrian", "Uruguaya" = "Uruguayan"),
    Birth = recode_factor(Birth, "aborto" = "Abortion", "intervencionista" = "Interventionist", "natural" = "Natural", "operatório" = "Operative"),
    MaternalOutcome = recode_factor(MaternalOutcome, "alta" = "Discharged", "morta" = "Death", "transferência" = "Hospital transferal"),
    FetalOutcome = recode_factor(FetalOutcome, "vivo" = "Live Birth", "morto" = "Stillbirth or Neonatal Death")
    )

# Save new summary data with recoded factor variables
saveRDS(ML_summary, file = here("data", "processed-data", "ML_summary_final.rds"))
```

@@

# Change observations from Portuguese to English
ML_linear <- ML_linear %>%
  mutate(
    ModifiedColor = recode_factor(ModifiedColor, "AfroDescent" = "Afro-Descent", "EuroDescent" = "Euro-Descent"),
    ModifiedNationality = recode_factor(ModifiedNationality, "European" = "European", "LatinAmerican" = "Latin American", "MiddleEastern" = "Middle Eastern", "Brazilian" = "Brazilian"),
    Color = recode_factor(Color, "Preta" = "Black", "Parda" = "Mixed Race", "Branca" = "White"),
    Nationality = recode_factor(Nationality, "Alema" = "German", "Argentina" = "Argentine", "Austriaca" = "Austrian", "Brasileira" = "Brazilian", "Espanhola" = "Spanish", "Francesa" = "French", "Italiana" = "Italian", "Paraguaya" = "Paraguayan", "Polaca" = "Polish", "Portuguesa" = "Portuguese", "Rumania" = "Romanian", "Russa" = "Russian", "Suiça" = "Swiss", "Siria" = "Syrian", "Uruguaya" = "Uruguayan"),
    Birth = recode_factor(Birth, "aborto" = "Abortion", "intervencionista" = "Interventionist", "natural" = "Natural", "operatório" = "Operative"),
    MaternalOutcome = recode_factor(MaternalOutcome, "alta" = "Discharged", "morta" = "Death", "transferência" = "Hospital transferal"),
    FetalOutcome = recode_factor(FetalOutcome, "vivo" = "Live Birth", "morto" = "Stillbirth or Neonatal Death")
    )

# Save new summary data with recoded factor variables
saveRDS(ML_linear, file = here("data", "processed-data", "ML_linear_final.rds"))
```

@@

For the previous code I worked on in relation to summary stats, see the `edacode-extra.R` file.
